master thesis

### 2018-03-24 Paper - Latex/Paper - Skizze - v4.tex
Paper - Results - Cross validation: refine text

  This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
    Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  The challenge here is the size of the tree and the little information about it. \\
  Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
    tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
    \cite{Poelen2014} which we used, we could only collect 25,992 parasitic and 34,879 
    free-living species. It follows that we have only $\approx 2.4~\%$ state information. \\
  So far, especially small scale studies have been carried out or highly manual. In this scale, it 
    requires different data sources to be interconnected. \\
  We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
    using the R package \textit{Castor} \cite{Louca2017}. \\
  Nevertheless, the results are convincing and even though purely computational approach which did 
    not include human experts input, results coincide with prior knowledge. Also regarding the number 
    of events, our estimates coincide with previous results by human experts, e.g. the study by 
    Weinstein and Kuris \cite{Weinstein2016}. \\

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{real data analysis}
    \todo{Evaluation section in methods?}
    \begin{itemize}
      \item Import tree
      \item Import interactions
      \item run castor algorithm / and others?
      \item interprete results (cross validation - leave 100 out)
    \end{itemize}
    We then evaluated the results in two ways. On the one hand, we have examined the results of some 
      subtrees for their credibility with the help of biological background knowledge. On the other 
      hand, we performed a leaf 100 out cross validation.

@@

  We then evaluated the results in two ways. On the one hand, we have examined the results of some 
    subtrees for their credibility with the help of biological background knowledge (subsection 
    \ref{sec:biological view}). On the other hand, we performed a leaf 100 out cross validation
    (Subsection \ref{sec:results cross evaluation}).

  % The next sections now guide you through the results of these different steps.

@@

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models were created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC multifurcation} 
      \end{table}
      % * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
      Within the every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (eg 
        family) were computationally too expensive to calculate. \\
      The model multifurc $\sim$ order * mean.height turns out to be the best of our models, whereby it 
        is possible that eg multifurc $\sim$ family * mean.height is better.

@@

      Next to the problem of the multifurcation of the tree is the less of interaction data we have for 
        the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 distinct free-living species and 47 332 
        distinct parasite species $\rightarrow$ unknown nodes 2,194,794 ($\approx 95.7~\%$). \\
      We found also 57,352 (not distinct!) source species and 809,993 (not distinct!) target
        species without OTT ids. Since we currently use only OTT ids, we could not use this information. \\
      % \todo{With this only $\approx 4.3~\%$ information in our leaf nodes are ...} \\
      \todo{discussion!!!} \\

@@

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BICs of unknown information}
        \label{table:BIC unknown information} 
      \end{table}
      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better. \\
      Again, the calculation of finer models (eg order or family) was too expensive. \\

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of the real data analysis created with Sankoff}
    This section is about evaluating the prediction of real data using the Sankoff method. \\
    It is divided into two subsections. \ref{sec:biological view}: The analysis of some subtrees using 
      biological background knowledge. \ref{sec:results cross evaluation}: Presentation of the results 
      of the leaf 100 out cross validation.

@@

      \begin{figure}[h!]
        \centering
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.85\textwidth]{Figures/ChordataParasites.png}
        \caption{Parasites of Phylum: Chordata. \\
          All parasite data of the chordata were mapped into a rough taxonomy (phylum, class, order, 
            family) in order to understand their distribution and affiliation. \\
          The internal nodes are the wanted taxa from OTL (with the addition of \# input parasites to 
            $\rightarrow$ \# predicted parasites). \\
          The leaf nodes are the input parasites (green) and the predicted parasites (white $\rightarrow$ 
            green).}
        \label{fig:ChordataParasites}
      \end{figure}

@@

        Weinstein and Kuris have been searching for origins of parasitism in Animalia \cite{Weinstein2016}. 
          They identified 223 parasitic origins: 223 in Metazoa $\supset$ 143 in Arthropoda $\supset$ 87 
          in Insecta. \\
        This has led us to count the origins and losses of parasitism in our investigation as well. \\
        We count only one origin / loss in a parent node with different children's nodes. \\
        Here we have encountered a problem: The Castor algorithm gives us probabilities for states. That 
          means there are also nodes with state like 0.3 or 0.5. The question is, how to count these state 
          changes. We decided to round the values.
        % \begin{lstlisting}[gobble=6]
        %   possible tags: 0, 0.333, 0.4, 0.5, 0.667, 0.75, 1
        %   rounded to:    0  0      0    0    1      1     1
        % \end{lstlisting}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Cross evaluation - leave 100 out} \label{sec:results cross evaluation}
      We ran the castor algorithm 100 times with leaving 100 randomized free-living or parasitic species 
        out of the input data to see how stable our result is. Of these 10,000 nodes, 9,238 were unique. 
        Of that, we predicted 9060 ($\approx 98.17~\%$) correctly and 169 ($\approx 1.82~\%$) wrongly, 
        with duplicate draws always having the same prediction. \\
      What is the best way to model this data? We again tested the influence of the taxa and the depth 
        of leaf nodes and calculated the BICs (Table: \ref{table:BIC cross validation}).

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r| }
            \hline
            \bfseries Model / Taxa &\bfseries  Kingdom & \bfseries Phylum & \bfseries Class \\% & Order \\
            \hline \hline
            correct predicted $\sim$ taxa & 117877 & \cellcolor{green!30}111466 & \cellcolor{green!45}108878 \\% & XXXXX \\
            \hline
            correct predicted $\sim$ taxa + depth & 117703 & \cellcolor{green!35}110513 & \cellcolor{green!45}108403 \\% & XXXXX \\
            \hline
            correct predicted $\sim$ taxa * depth & 117592 & \cellcolor{green!40}109827 & \cellcolor{green!50}107994 \\% & XXXXX \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{Residuals of cross validation prediction}
        \label{table:Residuals cross validation} 
      \end{table}

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class \\% & Order \\
            \hline \hline
            correct predicted $\sim$ taxa & 117936 & \cellcolor{green!40}112242 & \cellcolor{green!50}111733 \\% & XXXXX \\
            \hline
            correct predicted $\sim$ taxa + depth & 117776 & \cellcolor{green!50}111304 & \cellcolor{green!50}111273 \\% & XXXXX \\
            \hline
            correct predicted $\sim$ taxa * depth & 117709 & \cellcolor{green!50}111262 & \cellcolor{green!30}113135 \\% & XXXXX \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BICs of cross validation prediction}
        \label{table:BIC cross validation} 
      \end{table}

      Lower taxa than classes (eg order) were computationally too expensive to calculate. \\

      What could happen by removing a parasite or free-living of the list?
      \begin{itemize}
        \item It could be a specie, which don't exist in the tree leaf nodes. -> no effect
        \item It could be a specie, which exists in both lists. -> If it was a parasite, it is now 
          free-living, because we prefer parasites. Otherwise we have no effect again. (1053 are possible)
        \item Normal case: We loose information, because its a specie in our tree and we change it to a 
          leave node with no information.
      \end{itemize}

      Influence on the rest of the data:
      \begin{table}
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r|r| }
            \hline
            & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance ($\sigma^2$) & $\sigma$ \\
            \hline \hline
            \multirow{3}{*}{distance} & all   & 0 & 3587.70 & 224.96 & 313650.61 & 560.05 \\
            & leaf nodes                      & 0 & 3021.12 & 208.69 & 248103.38 & 498.10 \\
            & internal nodes                  & 0 & 566.58 & 16.28 & 4927.95 & 70.20 \\ \hline
            \multicolumn{2}{|c||}{changed tag} & 0 & 0 & 0 & 0 & 0 \\ \hline
            \multirow{3}{*}{lost} & all tags  & 100 & 100 & 100 & 0 & 0 \\
            & FL tags                         & 44 & 66 & 57.25 & 19.50 & 4.42 \\
            & P tags                          & 34 & 56 & 42.75 & 19.50 & 4.42 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics to Cross validation}
      \end{table}

      \begin{table}
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r| }
            \hline
            & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance \\
            \hline \hline
            \rowcolor{green!50}distance & all   & 1 & 2 & 1.33 & 0.33 \\
            \rowcolor{orange!50}&               & 0 & 3587.70 & 217.94 & 273760.68 \\
            \rowcolor{green!50}& leaf nodes     & 1 & 2 & 1.33 & 0.33 \\
            \rowcolor{orange!50}&               & 0 & 3021.12 & 202.57 & 209274.86 \\
            \rowcolor{green!50}& internal nodes& 0 & 0 & 0.00 & 0.00 \\
            \rowcolor{orange!50}&               & 0 & 566.58 & 15.37 & 5684.00 \\ \hline
            \rowcolor{green!50}lost & FL tags   & 44 & 49 & 46.67 & 6.33 \\
            \rowcolor{orange!50}&               & 51 & 66 & 57.95 & 13.47 \\
            \rowcolor{green!50}& P tags         & 51 & 56 & 53.33 & 6.33 \\
            \rowcolor{orange!50}&               & 34 & 49 & 42.05 & 13.47 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics to Cross validation \\
          - \colorbox{green!50}{Less free-livings - 3 examples} \\
          - \colorbox{orange!50}{Less parasites - 64 examples}}
      \end{table}

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Missing leaf state modelling - Resiudal tables}\label{sec:Residuals unknown information}
    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l|r|r|r| }
          \hline
          \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class \\
          \hline \hline
          multifurc $\sim$ taxa & 545740 & \cellcolor{green!10}499227 & \cellcolor{green!30}482265 \\
          \hline
          multifurc $\sim$ taxa + depth & 544789 & \cellcolor{green!10}493017 & \cellcolor{green!50}478998 \\
          \hline
          multifurc $\sim$ taxa * depth & 544062 & \cellcolor{green!30}488366 & \cellcolor{green!50}476382 \\
          \hline
        \end{tabular}
      \end{center}
      \caption{Residuals of unknown information}
    \end{table}

>>>>

  This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
    Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  The challenge here is the size of the tree and the little information about it. \\
  Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
    tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
    \cite{Poelen2014} which we used, we could only collect 25,992 parasitic and 34,879 free-living 
    species. It follows that we have $\approx 2.4~\%$ state information. \\
  So far, especially small scale studies have been carried out or highly manual. In this scale, it 
    requires different data sources to be interconnected. \\
  We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
    using the R package \textit{Castor} \cite{Louca2017}. \\
  Nevertheless, the results are convincing and even though purely computational approach which did 
    not include human experts input, results coincide with prior knowledge. Also regarding the number 
    of events, our estimates coincide with previous results by human experts, e.g. the study by 
    Weinstein and Kuris \cite{Weinstein2016}. \\

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{real data analysis}
    \todo{Evaluation section in methods?}
    \begin{itemize}
      \item Import tree
      \item Import interactions
      \item run castor algorithm / and others?
      \item interprete results (leave-100-out cross validation)
    \end{itemize}
    We then evaluated the results in two ways. On the one hand, we have examined the results of some 
      subtrees for their credibility with the help of biological background knowledge. On the other 
      hand, we performed a leave-100-out cross validation.

@@

  We then evaluated the results in two ways. On the one hand, we have examined the results of some 
    subtrees for their credibility with the help of biological background knowledge (subsection 
    \ref{sec:biological view}). On the other hand, we performed a leave-100-out cross validation
    (Subsection \ref{sec:results - cross validation}) and predicted approximately 98.17~\% states 
    correct.

@@

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models were created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC multifurcation} 
      \end{table}
      % * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
      Within the every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (eg 
        family) were computationally too expensive to calculate. \\
      The model \textit{multifurc $\sim$ order * mean.height} turns out to be the best of our models, 
        whereby it is possible that eg \textit{multifurc $\sim$ family * mean.height} is better.

@@

      Next to the problem of the multifurcation of the tree is the less of interaction data we have for 
        the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 distinct free-living species and 47 332 
        distinct parasite species for the whole tree of life. From the Eukaryota we could determine 
        25,992 and 34,879 species as parasites and free-living. With 2,293,463 leaf nodes we still have 
        about 97.34~\% unknown leaf nodes. \\
      We found also 57,352 (not distinct!) source species and 809,993 (not distinct!) target
        species without OTT ids. Since we currently use only OTT ids, we could not use this information. \\        
      \todo{discussion!!!} \\

@@

      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values unknown state information models. \\
          These models were created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC unknown information} 
      \end{table}
      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better. \\
      Again, the calculation of finer models (eg order or family) was too expensive. \\

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of the real data analysis created with Sankoff}
    This section is about evaluating the prediction of real data using the Sankoff method. \\
    It is divided into two subsections. \ref{sec:biological view}: The analysis of some subtrees using 
      biological background knowledge. \ref{sec:results - cross validation}: Presentation of the results 
      of the leave-100-out cross validation.

@@

      \begin{figure}
        \centering
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.85\textwidth]{Figures/ChordataParasites.png}
        \caption{Parasites of Phylum: Chordata. \\
          All parasite data of the chordata were mapped into a rough taxonomy (phylum, class, order, 
            family) in order to understand their distribution and affiliation. \\
          The internal nodes are the wanted taxa from OTL (with the addition of \# input parasites to 
            $\rightarrow$ \# predicted parasites). \\
          The leaf nodes are the input parasites (green) and the predicted parasites (white $\rightarrow$ 
            green).}
        \label{fig:ChordataParasites}
      \end{figure}

@@

        Weinstein and Kuris have been searching for origins of parasitism in Animalia \cite{Weinstein2016}. 
          They identified 223 parasitic origins: 223 in Metazoa $\supset$ 143 in Arthropoda $\supset$ 87 
          in Insecta. \\
        This has led us to count the origins and losses of parasitism in our investigation as well. \\
        We count only one origin / loss in a parent node with different children's nodes. \\
        Here we have encountered a problem: The Castor algorithm gives us probabilities for states. That 
          means there are also nodes with state like 0.3 or 0.5. The question is, how to count these state 
          changes. We decided to round the values.
        % \begin{lstlisting}[gobble=6]
        %   possible states: 0, 0.333, 0.4, 0.5, 0.667, 0.75, 1
        %   rounded to:    0  0      0    0    1      1     1
        % \end{lstlisting}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Leave-100-out cross validation} \label{sec:results - cross validation}
      For a further validation of our results, we carried out a leave-100-out cross validation. \\
      In order to achieve about 15\% of the 60,871 input node states with a validation, 
        we randomly left out 100 states 100 times. To make smaller amounts up to leave-one-out had too 
        much computational effort. \todo{warum 15\%?}\\

      Of these 10,000 nodes, 9,238 were unique. From the unique nodes, we predicted approximately 
        98.17~\% correct and thus 1.82~\% wrong, with duplicate draws always having the same prediction 
        (data not shown). \\

      We have again dealt with the question of how this data is best modeled. The influence of the taxa 
        (kingdom, phylum, class) and the depth of the leaf nodes was re-modeled and the BICs compared 
        (table \ref{table:BIC cross validation}). Lower taxa than classes (eg order) were 
        computationally too expensive to calculate. \\
      The Residuals are included in the appendix (table \ref{table:Residuals cross validation}) for the 
        sake of completeness; here too, due to the different complexity, they do not have much meaning 
        for the comparison of the models. \\

      \todo{table for these numbers?}
      % 100/(25992+34879)*10000 = 16.43
      % 100/(25992+34879)*9238 = 15.18
      % 10000 nodes 9238 unique
      % 9060 nodes $\approx 98.17~\%$ correctly predicted
      % 169 ($\approx 1.82~\%$) wrongly predicted

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class \\% & Order \\
            \hline \hline
            correct predicted $\sim$ taxa & 117936 & \cellcolor{green!40}112242 & \cellcolor{green!50}111733 \\% & XXXXX \\
            \hline
            correct predicted $\sim$ taxa + depth & 117776 & \cellcolor{green!50}111304 & \cellcolor{green!50}111273 \\% & XXXXX \\
            \hline
            correct predicted $\sim$ taxa * depth & 117709 & \cellcolor{green!50}111262 & \cellcolor{green!30}113135 \\% & XXXXX \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of cross validation prediction models. \\
          These models were created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values.}
        \label{table:BIC cross validation} 
      \end{table}

      The BIC values this time did not prove that the finest model is the best. Of our calculated models, 
       \textit{correct predicted $\sim$ phylum * depth} has the smallest value. \\

      We examined the influence of the omitted data on the prediction. On average, about twice as many 
        leaf nodes are predicted differently. \color{red}The variance is very high. This means that 
        we have a high degree of dispersion and thus a stochastic situation exists. It also describes 
        the width of the present probability function. The standard deviation is about five times as 
        high as the number of omitted nodes, so the variability is quintupled. \normalfont Table 
        \ref{table:statistics cross validation} shows these results. \\
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r|r| }
            \hline
            & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance ($\sigma^2$) & $\sigma$ \\
            \hline \hline
            \multirow{3}{*}{\bfseries distance} & \bfseries all     & 0 & 3587.70 & 224.96 & 313650.61 & 560.05 \\
            & \bfseries leaf nodes                                  & 0 & 3021.12 & 208.69 & 248103.38 & 498.10 \\
            & \bfseries internal nodes                              & 0 & 566.58 & 16.28 & 4927.95 & 70.20 \\ \hline
            \multirow{3}{*}{\bfseries lost} & \bfseries all states  & 100 & 100 & 100 & 0 & 0 \\
            & \bfseries FL states                                   & 44 & 66 & 57.25 & 19.50 & 4.42 \\
            & \bfseries P states                                    & 34 & 56 & 42.75 & 19.50 & 4.42 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics about the leave-100-out cross validation \\
          The distance between original and new states was calculated using the Euclidean metric. This 
            was summed over all states / all leaf node states and all internal node states. \\
          The lower half of the table describes the distribution of the 'lost' states between parasites 
            (P) and free-livings (FL).}
        \label{table:statistics cross validation}
      \end{table}
    %---------------------------------------------------------------------------------------------------

      % \begin{table}
      %   \begin{center}
      %     \begin{tabular}{ |cl||r|r|r|r| }
      %       \hline
      %       & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance \\
      %       \hline \hline
      %       \rowcolor{green!50}distance & all   & 1 & 2 & 1.33 & 0.33 \\
      %       \rowcolor{orange!50}&               & 0 & 3587.70 & 217.94 & 273760.68 \\
      %       \rowcolor{green!50}& leaf nodes     & 1 & 2 & 1.33 & 0.33 \\
      %       \rowcolor{orange!50}&               & 0 & 3021.12 & 202.57 & 209274.86 \\
      %       \rowcolor{green!50}& internal nodes& 0 & 0 & 0.00 & 0.00 \\
      %       \rowcolor{orange!50}&               & 0 & 566.58 & 15.37 & 5684.00 \\ \hline
      %       \rowcolor{green!50}lost & FL states & 44 & 49 & 46.67 & 6.33 \\
      %       \rowcolor{orange!50}&               & 51 & 66 & 57.95 & 13.47 \\
      %       \rowcolor{green!50}& P states       & 51 & 56 & 53.33 & 6.33 \\
      %       \rowcolor{orange!50}&               & 34 & 49 & 42.05 & 13.47 \\
      %       \hline
      %     \end{tabular}
      %   \end{center}
      %   \caption{Statistics to cross validation \\
      %     - \colorbox{green!50}{Less free-livings - 3 examples} \\
      %     - \colorbox{orange!50}{Less parasites - 64 examples}}
      % \end{table}

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Missing leaf state modelling - Resiudal table}\label{sec:Residuals unknown information}
    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l|r|r|r| }
          \hline
          \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class \\
          \hline \hline
          multifurc $\sim$ taxa & 545740 & \cellcolor{green!10}499227 & \cellcolor{green!30}482265 \\
          \hline
          multifurc $\sim$ taxa + depth & 544789 & \cellcolor{green!10}493017 & \cellcolor{green!50}478998 \\
          \hline
          multifurc $\sim$ taxa * depth & 544062 & \cellcolor{green!30}488366 & \cellcolor{green!50}476382 \\
          \hline
        \end{tabular}
      \end{center}
      \caption{Residuals of unknown information models \\
        These models were created with the R function \textit{glm()} and compared with the 
          \textit{anova()} function. This results in the listed residuals.}
    \end{table}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Cross validation prediction - Resiudal table}
    \begin{table}[h]
      \begin{center}
        \begin{tabular}{ |l|r|r|r| }
          \hline
          \bfseries Model / Taxa &\bfseries  Kingdom & \bfseries Phylum & \bfseries Class \\% & Order \\
          \hline \hline
          correct predicted $\sim$ taxa & 117877 & \cellcolor{green!30}111466 & \cellcolor{green!45}108878 \\% & XXXXX \\
          \hline
          correct predicted $\sim$ taxa + depth & 117703 & \cellcolor{green!35}110513 & \cellcolor{green!45}108403 \\% & XXXXX \\
          \hline
          correct predicted $\sim$ taxa * depth & 117592 & \cellcolor{green!40}109827 & \cellcolor{green!50}107994 \\% & XXXXX \\
          \hline
        \end{tabular} 
      \end{center}
      \caption{Residuals of cross validation prediction \\
        These models were created with the R function \textit{glm()} and compared with the 
          \textit{anova()} function. This results in the listed residuals.}
      \label{table:Residuals cross validation} 
    \end{table}
